---
title: "Land surface temperature from Modis"
author: "Faith Musili"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

## *Introduction*

*_Product: MOD11C3_*

- Provides monthly Land Surface Temperature and Emissivity (LST&E) values in a 0.05 degree (5,600 meters at the equator) latitude/longitude Climate Modeling Grid (CMG). 
- A CMG granule is a geographic grid with 7,200 columns and 3,600 rows representing the entire globe.
- The LST&E values in the MOD11C3 product are derived by compositing and averaging the values from the corresponding month of MOD11C1 daily files. 

*_Horizontal and vertical tiles_*

- Download all horizontal and vertical tiles for the world for future use.

*_Years_*

- MODIS TERRA satellite was launced in December 1999 and data was available from February 2000.
- images were downloaded from February 2000 to April 2019.

*_Collection version_*
 - Version 6 is the up to date version for this MODIS product.


```{r load required packages, eval=F}
library(MODISTools)
library(rgeos)
library(sp)
library(raster)
library(rgdal)
library(MODIS)
library(bitops)
library(RCurl)
library(XML)
library(weathermetrics)
library(raster)
library(gdalUtils)
```

### Data download
```{r download data , eval=F, echo=T}
MODISoptions(localArcPath="Download/")#### path to store data

getHdf(product="MOD11C3", ##### product
       begin="2000.01.01",##### the start day for download
       end="2000.02.01", ##### the last date for download
       tileH=0:35,##### input the horizontal tile as per the modis path and row  outline
       tileV=0:17,##### input the vertical tile as per the modis path and row outline
       extent=NULL, 
       collection=006,#### version to download
       quiet=FALSE, 
       wait=10)


```

### Convert hdf files to tif
```{r, eval=F, echo=T}
files <- list.files(pattern="\\.hdf$",full.names = FALSE,recursive = TRUE)

for (filename in files)
{
  sds <- get_subdatasets(filename)
  gdal_translate(sds[1], dst_dataset=paste0(substr(filename, 1, nchar(filename)-4) ,".tif"))
}
```

### Convert kelvin to degrees celcius
[guide](https://icess.eri.ucsb.edu/modis/LstUsrGuide/usrguide_month_cmg.html)

```{r, eval=F, echo=T}
tif.files <- list.files(pattern="\\.tif$",full.names = FALSE,recursive = TRUE)

### replace all  values to degrees celcius
for (i in 1:length(tif.files)){
  
  image<-raster(tif.files[i])## read files
  
  out.img<-kelvin.to.celsius(image, round = 1) ## convert the digital numbers from kelvin to degrees celcius
  
  writePath<-paste0(substr(tif.files[i], 1, nchar(tif.files[i])-4) ,"_degrees.tif")## output file path
  writeRaster(out.img,writePath,format="GTiff", overwrite=TRUE)## write file
}
```

### Stack all monthly images together into a stack
```{r, eval=F, echo=T}
## stack all months together
lstfinal<-list.files(pattern =".*_degrees.*\\.tif$" ,recursive = TRUE,full.names = FALSE)

lst.stack<-stack(lapply(lstfinal,raster))

timeseq<-paste0(sort(rep(2000:2019,12)),".",rep(month.abb,12))[-1]## order year and month

names(lst.stack)<-paste0("T",timeseq)## customize layer names
lstBrick<-brick(lst.stack)

tifoptions <- c("COMPRESS=DEFLATE", "PREDICTOR=2", "ZLEVEL=6")
writeRaster(lst.stack,"lstStack_monthly.tif",options = tifoptions, overwrite = TRUE)## write entire stack
```


## Crop Kenya lst 
### Monthly stack
```{r, eval=F, echo=T}
Kenyashp<-shapefile("shp/Kenya_shp.shp")
Kenya_extend<-extent(Kenyashp)
Kenyalst<-crop(lst.stack,Kenyashp)

writeRaster(Kenyalst,"KenyaLstStack.tif",format="GTiff", overwrite=TRUE)## write Kenya stack to memory
```

### Yearly stack (Stack all monthly images together into a yearly stack)
```{r, eval=F}
Kenyalst<-brick("KenyaLstStack.tif")
crs(Kenyalst)<-"+proj=longlat +datum=WGS84"

timeseq<-c(
  paste0(sort(rep(2000:2018,12)),".",rep(month.abb,12)),
  paste0(rep(2019,4),".",rep(month.abb[1:4],1)))[-1]## order year and month

names(Kenyalst)<-paste0("T",timeseq)## customize layer names

## remove the incomplete years
 Kenyalst<-Kenyalst[[c(12:227)]]## remove incomplete 2000 and 2019

 ranges<-data.frame(start=seq(1, 205, by=12),end=seq(12, 216, by=12))
 
 for (i in 1:nrow(ranges)) {

   start<-ranges$start[i]
   end<-ranges$end[i]
   
 img<-stackApply(Kenyalst[[start:end]],rep(1,12),fun=mean)
 
 outname<-paste0("Yearly_stack/",substr(names(Kenyalst[[start]]),2,5),"_lst.tif")
 
 writeRaster(img,outname,format="GTiff",overwite=TRUE)
 
 }
 
 lst_files<-list.files(path = "Yearly_stack/",pattern ="*.tif$" ,recursive = TRUE,full.names = TRUE)
 
 yrlstBrick<-brick(lapply(lst_files,raster))
 
 writeRaster(yrlstBrick,"Yearly_stack/Yearly_LST.tif",format="GTiff",overwite=TRUE)

```

